---
title: "Linking NaNDA With Your Data"
subtitle: "A walkthrough · {{< var session.date >}}"
author: "{{< var session.presenter >}} and {{< var session.facilitator >}}"
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: false
---


::: {.callout-tip}
## You don't need to run this

{{< var session.presenter >}} is running this code live. Every output below is already
rendered, so you can read along without installing anything. If you'd like to run it
yourself later, everything is in the [materials ZIP]({{< var site.zip >}}).

**Work in Stata or Python?** Every code block has three tabs. The live session and the
rendered output are R; the Stata and Python tabs show the same step in your language.
Pick a tab once and the whole page follows.
:::

---

# 0. Setup {#sec-setup}

*Before we start — {{< var timing.open >}} minutes*


::: {.panel-tabset group="language"}
## R

```{r}
#| label: setup
#| message: false

library(tidyverse)   # dplyr for the joins, readr for the CSVs
library(readxl)      # the crosswalk is an .xlsx; installs with tidyverse, loads separately
library(here)        # file paths that work wherever this folder sits on your machine
library(janitor)     # tabyl(), for frequency tables you can read
```

## Stata

```{.stata}
* Run this from the workshop folder (the one holding workshop.Rproj), so the
* relative paths below resolve: File > Change Working Directory, or `cd`.
version 16
clear all
set more off

* One tempfile per table the R version keeps in memory. Stata holds one
* dataset at a time, so each step saves and the next step merges it back.
tempfile mydata socialservices parks ses2010 ses2020 xwalk mydata_zcta10 mydata_nanda
```

## Python

```{.python}
import pandas as pd            # data frames, read_csv / read_stata / merge
from pathlib import Path       # file paths that work wherever this folder sits

# The workshop folder: the one holding workshop.Rproj. Adjust if you moved the script.
ROOT = Path(__file__).resolve().parents[1] if "__file__" in globals() else Path.cwd()
DATA = ROOT / "data"
```

:::

::: {.panel-tabset group="language"}
## R

```{r}
#| label: read-data
#| message: false

# Our own data. `zip` is a label, not a quantity: say so, or read_csv guesses
# "number" and 03042 silently becomes 3042.
mydata <- read_csv(
  here("data", "synthetic_data_v20260924.csv"),
  col_types = cols(zip = col_character())
)

# NaNDA: Social Services, ZCTA 2010, 1990-2022. Same rule for the ZCTA code.
socialservices_zcta10 <- read_csv(
  here("data", "nanda", "nanda_socials_Zcta10_1990-2022_01.csv"),
  col_types = cols(zcta10 = col_character())
)

# Parks and Socioeconomic Status come from ICPSR as .rda files. load() drops an
# object into your session under a name ICPSR picked, and returns that name — so
# print it, then rename to something you'll recognise three screens from now.
print(load(here("data", "nanda", "38586-0002-Data.rda")))
parks2022_zcta10 <- da38586.0002
rm(da38586.0002)

print(load(here("data", "nanda", "38528-0003-Data.rda")))   # SES, ZCTA 2010, 2008-2017
ses2008_2017_zcta10 <- da38528.0003
rm(da38528.0003)

print(load(here("data", "nanda", "38528-0008-Data.rda")))   # SES, ZCTA 2020, 2018-2022
ses2018_2022_zcta20 <- da38528.0008
rm(da38528.0008)
```

## Stata

```{.stata}
* Our own data. Stata guesses "number" for a column of digits, and 03042 comes
* in as 3042. Check what it guessed and pad back to five characters if so.
import delimited using "data/synthetic_data_v20260924.csv", varnames(1) clear
capture confirm string variable zip
if _rc tostring zip, replace format(%05.0f)
save `mydata'

* NaNDA: Social Services, ZCTA 2010, 1990-2022. Same rule for the ZCTA code.
import delimited using "data/nanda/nanda_socials_Zcta10_1990-2022_01.csv", varnames(1) clear
capture confirm string variable zcta10
if _rc tostring zcta10, replace format(%05.0f)
save `socialservices'

* Parks and Socioeconomic Status: take the Stata format from ICPSR. Variable
* names arrive in upper case, as the ICPSR codebook lists them.
use "data/nanda/38586-0002-Data.dta", clear          // Parks, ZCTA, 2022
capture confirm string variable ZCTA19
if _rc tostring ZCTA19, replace format(%05.0f)
save `parks'

use "data/nanda/38528-0003-Data.dta", clear          // SES, ZCTA 2010, 2008-2017
capture confirm string variable ZCTA10
if _rc tostring ZCTA10, replace format(%05.0f)
save `ses2010'

use "data/nanda/38528-0008-Data.dta", clear          // SES, ZCTA 2020, 2018-2022
capture confirm string variable ZCTA20
if _rc tostring ZCTA20, replace format(%05.0f)
save `ses2020'
```

## Python

```{.python}
# Our own data. `zip` is a label, not a quantity: say so, or pandas guesses
# "number" and 03042 silently becomes 3042.
mydata = pd.read_csv(DATA / "synthetic_data_v20260924.csv", dtype={"zip": str})

# NaNDA: Social Services, ZCTA 2010, 1990-2022. Same rule for the ZCTA code.
socialservices_zcta10 = pd.read_csv(
    DATA / "nanda" / "nanda_socials_Zcta10_1990-2022_01.csv", dtype={"zcta10": str}
)

# Parks and Socioeconomic Status: take the Stata format from ICPSR; pandas reads
# it without extra packages. Force every ZCTA code to five-character text.
def zcta_text(df, col):
    df[col] = df[col].astype(str).str.replace(r"\.0$", "", regex=True).str.zfill(5)
    return df

parks2022_zcta10    = zcta_text(pd.read_stata(DATA / "nanda" / "38586-0002-Data.dta"), "ZCTA19")
ses2008_2017_zcta10 = zcta_text(pd.read_stata(DATA / "nanda" / "38528-0003-Data.dta"), "ZCTA10")
ses2018_2022_zcta20 = zcta_text(pd.read_stata(DATA / "nanda" / "38528-0008-Data.dta"), "ZCTA20")
```

:::

::: {.callout-warning}
## Read ZIP codes as text, always

ZIP codes are labels, not quantities: read them as numbers and `03042` becomes `3042`,
nothing warns you, and every join for that row fails. In practice that is all of New
England, New Jersey, and Puerto Rico quietly dropping out of your analysis.
:::

::: {.panel-tabset group="language"}
## R

```{r}
#| label: glimpse-data

glimpse(mydata)
```

## Stata

```{.stata}
use `mydata', clear
describe
list in 1/5
```

## Python

```{.python}
mydata.info()
mydata.head()
```

:::

---

# 1. What identifiers do you actually have? {#sec-identifiers}

*About {{< var timing.sec1 >}} minutes*

Most survey and clinical datasets do not have geography. They have a string: an
address, a city, a ZIP code, collected for mailing or billing and never checked for the
second job it is now being asked to do.

That job is to place each person in a neighborhood that some other dataset has
measured. Whether the string can do it depends on two things: the unit the other
dataset is published at, and what it takes to get from what you have to that unit.

So the question before any merge is not "does my data have location?" It is: which
identifier do I actually have, how complete is it, and what is it an identifier *of*? A
ZIP code is not a place in the way a Census tract is a place, and an address is not an
identifier at all until something turns it into one.

If your data already carry Census tract or ZCTA codes (ZIP Code Tabulation Areas, the
Census Bureau's version of ZIP codes) from an earlier geocoding step, you are ahead,
but not exempt. Those codes have a vintage, meaning the census year whose boundaries
they follow, and the vintage question in Section 3 applies with full force.


## What's in the synthetic dataset

| Variable | Type | Notes |
|---|---|---|
| `sampleid` | numeric | 2,004 unique — one row per person |
| `year` | numeric | {{< var synthetic.years >}} |
| `address`, `city`, `state`, `zip` | character | `zip` missing for ~{{< var synthetic.zip_missing_pct >}} of rows |
| `social_interactions`, `physical_activity` | 1–5 | |
| `hot_meal_days` | 1–7 | |
| `loneliness` | continuous | |
| `own_pet` | 0/1 | |

## What NaNDA is published at

NaNDA describes neighborhoods, not people. Every measure is a count, a density, or a
rate for a geographic unit in a given year. There are no individual records anywhere in
the archive, which is why the linking is on you: NaNDA cannot know where your
participants live.

Today we use two datasets:

- **{{< var nanda.social_services_name >}}** ({{< var nanda.social_services_years >}}),
  <{{< var nanda.social_services_doi >}}>
- **{{< var nanda.parks_name >}}** ({{< var nanda.parks_years >}}),
  <{{< var nanda.parks_doi >}}>

Both are released at the Census tract and the ZCTA level. We use the ZCTA files, because
that is the closest thing to what the synthetic data can offer.

Which is the point to sit with: our file has ZIP codes, and NaNDA has ZCTAs. They look
alike, five digits and usually the same five digits, and they are not the same thing.
The next section is about that gap.

::: {.panel-tabset group="language"}
## R

```{r}
#| label: inspect-identifiers

# How many rows carry a ZIP we can actually look something up with?
mydata %>%
  summarise(
    rows        = n(),
    has_zip     = sum(!is.na(zip)),
    missing_zip = sum(is.na(zip)),
    pct_missing = round(100 * mean(is.na(zip)), 1)
  )
```

## Stata

```{.stata}
* How many rows carry a ZIP we can actually look something up with?
use `mydata', clear
generate byte missing_zip = missing(zip)
summarize missing_zip
display "rows = " r(N) ", missing ZIP = " r(sum) ", pct missing = " %4.1f 100 * r(mean)
```

## Python

```{.python}
# How many rows carry a ZIP we can actually look something up with?
pd.Series({
    "rows":        len(mydata),
    "has_zip":     mydata["zip"].notna().sum(),
    "missing_zip": mydata["zip"].isna().sum(),
    "pct_missing": round(100 * mydata["zip"].isna().mean(), 1),
})
```

:::


::: {.panel-tabset group="language"}
## R

```{r}
#| label: inspect-address

# address, city and zip are each incomplete — and not for the same rows. A row
# with no ZIP may still have an address you could geocode, and the other way
# round, so "how much geography do I have" depends on which route you take.
mydata %>%
  count(
    address_missing = is.na(address),
    city_missing    = is.na(city),
    zip_missing     = is.na(zip)
  ) %>%
  arrange(desc(n))
```

## Stata

```{.stata}
* address, city and zip are each incomplete, and not for the same rows. A row
* with no ZIP may still have an address you could geocode, and the other way
* round, so "how much geography do I have" depends on which route you take.
use `mydata', clear
generate byte address_missing = missing(address)
generate byte city_missing    = missing(city)
generate byte zip_missing     = missing(zip)
contract address_missing city_missing zip_missing, freq(n)
gsort -n
list, clean noobs
```

## Python

```{.python}
# address, city and zip are each incomplete, and not for the same rows. A row
# with no ZIP may still have an address you could geocode, and the other way
# round, so "how much geography do I have" depends on which route you take.
(mydata
   .assign(address_missing=mydata["address"].isna(),
           city_missing=mydata["city"].isna(),
           zip_missing=mydata["zip"].isna())
   .value_counts(["address_missing", "city_missing", "zip_missing"])
   .rename("n")
   .reset_index()
   .sort_values("n", ascending=False))
```

:::


::: {.callout-note}
## Aside: what if you have addresses?

Street addresses are a different workflow, and we are not doing it today. You would
geocode each address to coordinates, assign those coordinates to a Census tract, and use
the tract-level NaNDA file, with failure modes of its own: addresses that will not
parse, rooftop versus street-segment precision, and tract boundaries that also moved in
2020. The Census Bureau's batch geocoder (<https://geocoding.geo.census.gov/>) is free,
needs no API key, and returns the tract code directly, which skips the spatial join
entirely. If your data have addresses, that tool plus Section 3 is most of what you
need.
:::

---

# 2. ZIP is not ZCTA {#sec-crosswalk}

*About {{< var timing.sec2 >}} minutes*

A ZIP code is a Postal Service artifact. It names a set of delivery routes, and it
exists to get mail to a door, not to describe a place. Some ZIP codes are a single
building. Some are a bank of post office boxes with no residents at all. Some cross
county and even state lines, and the Postal Service redraws them whenever the mail
requires, without telling anyone doing research.

A ZCTA, a ZIP Code Tabulation Area, is the Census Bureau's answer to the fact that
researchers use ZIP codes anyway. For each decennial census, the Bureau takes every
census block, looks at the ZIP code most of its addresses carry, and assigns the block
to that ZIP's ZCTA. The result is a set of areas with the same five-digit names as ZIP
codes that tile the country and can be mapped, measured, and published at. NaNDA's ZCTA
files are built on these.

Most of the time a ZIP code and the ZCTA of the same name cover roughly the same people.
"Most of the time" is doing a lot of work in that sentence. ZIP codes with no ZCTA, such
as post office boxes and single-site ZIPs, have nowhere to go. ZIPs whose deliveries were
mostly absorbed into a neighboring ZCTA are assigned there, under a different number. So
the step from ZIP to ZCTA is a lookup against a crosswalk, not a rename of the column.
And the crosswalk has a vintage of its own.

## The crosswalk we're using

- **{{< var crosswalk.primary_name >}}** ({{< var crosswalk.primary_year >}}) —
  <{{< var crosswalk.primary_url >}}>
- NaNDA's own merge code: {{< var crosswalk.merge_code_name >}}
  ({{< var crosswalk.merge_code_year >}}), {{< var crosswalk.merge_code_doi >}}

Use NaNDA's published merge code rather than a snippet copied from a slide: it is
citable, it is maintained, and when something needs fixing the fix lands in one place
you can find again.

::: {.panel-tabset group="language"}
## R

```{r}
#| label: read-crosswalk

# Every column as text, for the same reason as `zip` above. Then pad back to five
# characters: whatever stripped a leading zero upstream did it before we got here.
zipzctaxwalk2019 <- read_excel(
  here("data", "zip_to_zcta_2019.xlsx"),
  col_types = "text"
) %>%
  mutate(
    ZIP_CODE = str_pad(ZIP_CODE, 5, pad = "0"),
    ZCTA     = str_pad(ZCTA,     5, pad = "0")
  )

# zip_join_type is the crosswalk telling you how each ZIP got its ZCTA — or that
# it has none. Read this before you join, not after something looks wrong.
tabyl(zipzctaxwalk2019, zip_join_type)
```

## Stata

```{.stata}
* Every column as text, for the same reason as `zip` above. Then pad back to
* five characters: whatever stripped a leading zero upstream did it before we
* got here.
import excel using "data/zip_to_zcta_2019.xlsx", firstrow allstring clear
replace ZIP_CODE = substr("00000", 1, 5 - strlen(ZIP_CODE)) + ZIP_CODE if strlen(ZIP_CODE) < 5
replace ZCTA     = substr("00000", 1, 5 - strlen(ZCTA))     + ZCTA     if strlen(ZCTA)     < 5
keep ZIP_CODE ZCTA zip_join_type
save `xwalk'

* zip_join_type is the crosswalk telling you how each ZIP got its ZCTA, or that
* it has none. Read this before you join, not after something looks wrong.
tabulate zip_join_type
```

## Python

```{.python}
# Every column as text, for the same reason as `zip` above. Then pad back to
# five characters: whatever stripped a leading zero upstream did it before we
# got here. (read_excel needs the openpyxl package: pip install openpyxl)
zipzctaxwalk2019 = pd.read_excel(DATA / "zip_to_zcta_2019.xlsx", dtype=str)
zipzctaxwalk2019["ZIP_CODE"] = zipzctaxwalk2019["ZIP_CODE"].str.zfill(5)
zipzctaxwalk2019["ZCTA"]     = zipzctaxwalk2019["ZCTA"].str.zfill(5)

# zip_join_type is the crosswalk telling you how each ZIP got its ZCTA, or that
# it has none. Read this before you join, not after something looks wrong.
zipzctaxwalk2019["zip_join_type"].value_counts(dropna=False)
```

:::

::: {.panel-tabset group="language"}
## R

```{r}
#| label: crosswalk-join

mydata_zcta10 <- mydata %>%
  left_join(
    zipzctaxwalk2019 %>% select(ZIP_CODE, ZCTA, zip_join_type),
    by = c("zip" = "ZIP_CODE")
  ) %>%
  rename(zcta10 = ZCTA, zip_join_type10 = zip_join_type)

# A left join keeps every row of the left-hand table. Check it; don't assume it.
# If this number grew, the crosswalk has more than one row per ZIP and you have
# quietly duplicated people.
c(rows_before = nrow(mydata), rows_after = nrow(mydata_zcta10))

# Two different failures, worth keeping apart. A row with no ZIP never had
# anything to look up. A row whose ZIP is missing from the crosswalk had one, and
# it led nowhere — PO-box ZIPs, single-building ZIPs, ZIPs the USPS has retired.
mydata_zcta10 %>%
  mutate(zcta_status = case_when(
    is.na(zip)    ~ "no ZIP",
    is.na(zcta10) ~ "ZIP not in crosswalk",
    TRUE          ~ "has a ZCTA"
  )) %>%
  tabyl(zcta_status)
```

## Stata

```{.stata}
use `xwalk', clear
rename (ZIP_CODE ZCTA zip_join_type) (zip zcta10 zip_join_type10)
save `xwalk', replace

use `mydata', clear
count
local rows_before = r(N)

* m:1 says "many of mine to one of theirs". If the crosswalk has two rows for
* one ZIP, Stata stops here instead of quietly duplicating people. keep(master
* match) is the left join: every row of ours stays, matched or not.
merge m:1 zip using `xwalk', keep(master match)
count
display "rows before = `rows_before', rows after = " r(N)

* Two different failures, worth keeping apart. A row with no ZIP never had
* anything to look up. A row whose ZIP is missing from the crosswalk had one,
* and it led nowhere: PO-box ZIPs, single-building ZIPs, retired ZIPs.
generate str zcta_status = "has a ZCTA"
replace  zcta_status = "ZIP not in crosswalk" if _merge == 1
replace  zcta_status = "no ZIP"               if missing(zip)
tabulate zcta_status
drop _merge
save `mydata_zcta10'
```

## Python

```{.python}
mydata_zcta10 = (
    mydata.merge(
        zipzctaxwalk2019[["ZIP_CODE", "ZCTA", "zip_join_type"]],
        how="left", left_on="zip", right_on="ZIP_CODE",
    )
    .drop(columns="ZIP_CODE")
    .rename(columns={"ZCTA": "zcta10", "zip_join_type": "zip_join_type10"})
)

# A left join keeps every row of the left-hand table. Check it; don't assume it.
# If this number grew, the crosswalk has more than one row per ZIP and you have
# quietly duplicated people.
{"rows_before": len(mydata), "rows_after": len(mydata_zcta10)}

# Two different failures, worth keeping apart. A row with no ZIP never had
# anything to look up. A row whose ZIP is missing from the crosswalk had one,
# and it led nowhere: PO-box ZIPs, single-building ZIPs, retired ZIPs.
import numpy as np
mydata_zcta10.assign(zcta_status=np.select(
    [mydata_zcta10["zip"].isna(), mydata_zcta10["zcta10"].isna()],
    ["no ZIP", "ZIP not in crosswalk"], default="has a ZCTA",
))["zcta_status"].value_counts()
```

:::

::: {.callout-important}
## Which ZCTA vintage does your crosswalk target?

The {{< var crosswalk.primary_year >}} release of the {{< var crosswalk.primary_name >}}
targets **{{< var crosswalk.primary_zcta_vintage >}} ZCTAs**. That makes it the right
partner for the 2010-vintage NaNDA file and the wrong one for the 2020-vintage file. In
the next section we join to both anyway, on purpose, because nothing in the code will
tell you which of the two joins was mismatched. When you do this with your own data,
read the crosswalk's documentation for its target vintage before you read anything else.


:::

---

# 3. Which vintage? Joining, and checking the join {#sec-join}

*About {{< var timing.sec3 >}} minutes — **this is the section that matters***

The join is three lines. Nobody gets the three lines wrong. The software sees to that,
and if a column name is misspelled it says so. What people get wrong happens on either
side of those lines and makes no sound at all.

The first silent error is vintage. ZCTA boundaries were redrawn for 2020. Most codes
survived with the same five digits and slightly different edges; some appeared, some
vanished, some were split. A 2020 code joined against a 2010 file will usually find a
row, and the row will usually describe approximately the same place. Approximately.
Your regression will not carry a footnote saying which cases were approximate.

The second silent error is never looking at what fell out. A left join keeps every row
of your data and fills NaNDA's columns with `NA` where nothing matched. An inner join
drops those rows. Either way the analysis proceeds, and unless you count, you will not
know how many people you lost or who they were. Stata's `merge` at least hands you
`_merge` and reports it. R's join functions (from the dplyr package) report nothing, so
the counting has to be built in by hand.

The rest of this section is therefore mostly diagnostics, and that is the point. The
join is the cheap part.

## Carrying Social Services forward

Social Services ends in 2022; our data runs to 2025. Before joining, we give every ZCTA
rows through 2025 and carry its most recent values into them. That is a decision, not a
fix: the 2023 to 2025 rows will match, and what they match is 2022's numbers. Write the
decision down in your methods, because nothing in the output will.

::: {.panel-tabset group="language"}
## R

```{r}
#| label: carry-forward

# complete() adds the missing ZCTA-year rows; fill() carries each ZCTA's last
# value down into them. fill() does not stop at 2022: any value missing inside
# 1990-2022 (population, and the densities where population is zero) also gets
# the year before. Know that it happens.
socialservices_zcta10 <- socialservices_zcta10 %>%
  complete(zcta10, year = 1990:2025) %>%
  arrange(zcta10, year) %>%
  group_by(zcta10) %>%
  fill(everything(), .direction = "down") %>%
  ungroup()

# Every year from 2022 on should now carry the same values.
tabyl(socialservices_zcta10, year) %>% filter(year >= 2020)
```

## Stata

```{.stata}
* Stata has no complete(); every ZCTA here has a 2022 row, so make three copies
* of it and number them 2023-2025. Then carry each ZCTA's last value down into
* anything still missing. Like R's fill(), that also reaches values missing
* inside 1990-2022 (population, and densities where population is zero).
use `socialservices', clear
expand 4 if year == 2022
bysort zcta10 year: replace year = 2021 + _n if year == 2022
sort zcta10 year
foreach v of varlist _all {
    if !inlist("`v'", "zcta10", "year") {
        by zcta10: replace `v' = `v'[_n-1] if missing(`v')
    }
}
save `socialservices', replace

* Every year from 2022 on should now carry the same values.
tabulate year if year >= 2020
```

## Python

```{.python}
# pandas has no complete(); every ZCTA here has a 2022 row, so copy it to
# 2023-2025. Then carry each ZCTA's last value down into anything still missing.
# Like R's fill(), that also reaches values missing inside 1990-2022
# (population, and densities where population is zero).
ss_2022 = socialservices_zcta10[socialservices_zcta10["year"] == 2022]
socialservices_zcta10 = (
    pd.concat([socialservices_zcta10] + [ss_2022.assign(year=y) for y in (2023, 2024, 2025)])
    .sort_values(["zcta10", "year"])
    .reset_index(drop=True)
)
cols = socialservices_zcta10.columns.drop("zcta10")
socialservices_zcta10[cols] = socialservices_zcta10.groupby("zcta10")[cols].ffill()

# Every year from 2022 on should now carry the same values.
socialservices_zcta10.loc[socialservices_zcta10["year"] >= 2020, "year"].value_counts().sort_index()
```

:::

## The join

::: {.panel-tabset group="language"}
## R

```{r}
#| label: join-nanda

mydata_nanda <- mydata_zcta10 %>%
  # Social Services is longitudinal, so the key is ZCTA *and* year.
  # `in_nanda` is added before the join: it is TRUE on every row of the NaNDA
  # side, so after the join it is TRUE where something matched and NA where
  # nothing did. That is the diagnostic Stata hands you as `_merge` and dplyr
  # does not — if you want it, you build it.
  left_join(
    socialservices_zcta10 %>%
      select(zcta10, year, totpop, aland10,
             count_totindivfamilyservices, den_totindivfamilyservices,
             count_childyouthservices,     den_childyouthservices,
             count_elderservices,          den_elderservices) %>%
      mutate(in_nanda = TRUE),
    by = c("zcta10", "year")
  ) %>%
  mutate(in_nanda = coalesce(in_nanda, FALSE)) %>%
  # Parks is a single 2022 snapshot, so it joins on ZCTA alone and every year of
  # a person's records gets the same value. That is an assumption — that park
  # provision held still across the study period — not a free lunch.
  left_join(
    parks2022_zcta10 %>%
      select(ZCTA19, ANY_OPEN_PARK, COUNT_OPEN_PARKS_TC10, PROP_PARK_AREA_ZCTA),
    by = c("zcta10" = "ZCTA19")
  )

c(rows_before = nrow(mydata_zcta10), rows_after = nrow(mydata_nanda))
```

## Stata

```{.stata}
use `socialservices', clear
keep zcta10 year totpop aland10 ///
     count_totindivfamilyservices den_totindivfamilyservices ///
     count_childyouthservices     den_childyouthservices ///
     count_elderservices          den_elderservices
save `socialservices', replace

use `mydata_zcta10', clear
count
local rows_before = r(N)

* Social Services is longitudinal, so the key is ZCTA *and* year. Stata's
* _merge is the diagnostic the R version has to build by hand: 3 = matched,
* 1 = ours only. Keep it as in_nanda.
merge m:1 zcta10 year using `socialservices', keep(master match)
generate byte in_nanda = (_merge == 3)
drop _merge

* Parks is a single 2022 snapshot, so it joins on ZCTA alone and every year of
* a person's records gets the same value. That is an assumption, that park
* provision held still across the study period, not a free lunch.
preserve
    use `parks', clear
    keep ZCTA19 ANY_OPEN_PARK COUNT_OPEN_PARKS_TC10 PROP_PARK_AREA_ZCTA
    rename ZCTA19 zcta10
    save `parks', replace
restore
merge m:1 zcta10 using `parks', keep(master match) nogenerate

count
display "rows before = `rows_before', rows after = " r(N)
save `mydata_nanda'
```

## Python

```{.python}
# Social Services is longitudinal, so the key is ZCTA *and* year.
# merge(indicator=True) is the diagnostic Stata hands you as _merge: "both"
# where something matched, "left_only" where nothing did. Keep it as in_nanda.
mydata_nanda = (
    mydata_zcta10.merge(
        socialservices_zcta10[[
            "zcta10", "year", "totpop", "aland10",
            "count_totindivfamilyservices", "den_totindivfamilyservices",
            "count_childyouthservices",     "den_childyouthservices",
            "count_elderservices",          "den_elderservices",
        ]],
        how="left", on=["zcta10", "year"], indicator=True,
    )
    .assign(in_nanda=lambda d: d["_merge"].eq("both"))
    .drop(columns="_merge")
    # Parks is a single 2022 snapshot, so it joins on ZCTA alone and every year
    # of a person's records gets the same value. That is an assumption, that
    # park provision held still across the study period, not a free lunch.
    .merge(
        parks2022_zcta10[["ZCTA19", "ANY_OPEN_PARK", "COUNT_OPEN_PARKS_TC10", "PROP_PARK_AREA_ZCTA"]],
        how="left", left_on="zcta10", right_on="ZCTA19",
    )
    .drop(columns="ZCTA19")
)

{"rows_before": len(mydata_zcta10), "rows_after": len(mydata_nanda)}
```

:::

## Socioeconomic status: two files, split by year

NaNDA's SES measures are five-year ACS estimates, so each file's value stands for a
stretch of years. We split our rows at 2018. Rows through 2017 join the 2008–2017 file,
published at 2010 ZCTAs. Rows from 2018 on join the 2018–2022 file, published at 2020
ZCTAs, and carry its values forward through 2025. Then the two halves are stacked back
into one dataset.

::: {.panel-tabset group="language"}
## R

```{r}
#| label: ses-split

# The two files name the affluence measure differently; give them one name
# before stacking.
mydata_nanda_le2017 <- mydata_nanda %>%
  filter(year <= 2017) %>%
  left_join(
    ses2008_2017_zcta10 %>%
      select(ZCTA10, AFFLUENCE13_17) %>%
      rename(AFFLUENCE = AFFLUENCE13_17),
    by = c("zcta10" = "ZCTA10")
  )

# Note: ZIP codes and ZCTAs can change geographically over time. These are
# 2010-vintage codes from the crosswalk, joined to a file drawn on 2020 ZCTAs.
mydata_nanda_ge2018 <- mydata_nanda %>%
  filter(year >= 2018) %>%
  left_join(
    ses2018_2022_zcta20 %>% select(ZCTA20, AFFLUENCE),
    by = c("zcta10" = "ZCTA20")
  )

mydata_nanda <- bind_rows(mydata_nanda_le2017, mydata_nanda_ge2018)

c(rows = nrow(mydata_nanda), with_affluence = sum(!is.na(mydata_nanda$AFFLUENCE)))
```

## Stata

```{.stata}
* The two files name the affluence measure differently; give them one name
* before stacking.
use `ses2010', clear
keep ZCTA10 AFFLUENCE13_17
rename (ZCTA10 AFFLUENCE13_17) (zcta10 AFFLUENCE)
save `ses2010', replace

use `ses2020', clear
keep ZCTA20 AFFLUENCE
rename ZCTA20 zcta10
save `ses2020', replace

use `mydata_nanda', clear
keep if year <= 2017
merge m:1 zcta10 using `ses2010', keep(master match) nogenerate
tempfile le2017
save `le2017'

* Note: ZIP codes and ZCTAs can change geographically over time. These are
* 2010-vintage codes from the crosswalk, merged to a file drawn on 2020 ZCTAs.
use `mydata_nanda', clear
keep if year >= 2018
merge m:1 zcta10 using `ses2020', keep(master match) nogenerate

append using `le2017'
count if !missing(AFFLUENCE)
display "rows = " _N ", with affluence = " r(N)
save `mydata_nanda', replace
```

## Python

```{.python}
# The two files name the affluence measure differently; give them one name
# before stacking.
ses2010 = ses2008_2017_zcta10[["ZCTA10", "AFFLUENCE13_17"]].rename(
    columns={"ZCTA10": "zcta10", "AFFLUENCE13_17": "AFFLUENCE"})
ses2020 = ses2018_2022_zcta20[["ZCTA20", "AFFLUENCE"]].rename(columns={"ZCTA20": "zcta10"})

# Note: ZIP codes and ZCTAs can change geographically over time. The 2018-on
# half joins 2010-vintage codes from the crosswalk to a file drawn on 2020 ZCTAs.
mydata_nanda = pd.concat([
    mydata_nanda[mydata_nanda["year"] <= 2017].merge(ses2010, how="left", on="zcta10"),
    mydata_nanda[mydata_nanda["year"] >= 2018].merge(ses2020, how="left", on="zcta10"),
], ignore_index=True)

{"rows": len(mydata_nanda), "with_affluence": mydata_nanda["AFFLUENCE"].notna().sum()}
```

:::

## 2010 vs 2020 ZCTAs, side by side

The note in that code is doing real work. The crosswalk gave us 2010 ZCTA codes, and
from 2018 on we joined them to a file drawn on 2020 ZCTAs. To see what that costs,
take the stack apart. What follows is not a demonstration of a mistake. It is the same
join run twice, against the 2010-vintage NaNDA file and the 2020-vintage file, using the
same crosswalked ZCTA codes, across every row. Both runs succeed. Both match the same rows, at the same rate, because most ZCTAs
kept their five digits from one census to the next. What differs is the value each run
attaches to those rows: the same people get a different mean affluence depending on
which file was open. Nothing in either output identifies which run paired codes with
the geography they were drawn for. That can only be answered from outside the code, by
knowing what vintage each input carries.

::: {.panel-tabset group="language"}
## R

```{r}
#| label: vintage-2010

# Socioeconomic Status, ZCTA 2010 file (ICPSR 38528, DS0003). Its affluence
# measure is named for the ACS years behind it, so rename it to a common name
# now — the two files we are comparing call the same construct different things.
vintage_2010 <- mydata_zcta10 %>%
  left_join(
    ses2008_2017_zcta10 %>%
      select(ZCTA10, AFFLUENCE13_17) %>%
      rename(AFFLUENCE = AFFLUENCE13_17),
    by = c("zcta10" = "ZCTA10")
  )

vintage_2010 %>%
  summarise(
    rows           = n(),
    matched        = sum(!is.na(AFFLUENCE)),
    match_rate_pct = round(100 * mean(!is.na(AFFLUENCE)), 1),
    mean_affluence = round(mean(AFFLUENCE, na.rm = TRUE), 3)
  )
```

## Stata

```{.stata}
* Socioeconomic Status, ZCTA 2010 file (ICPSR 38528, DS0003), already trimmed
* and renamed above, against every row rather than just the rows through 2017.
use `mydata_zcta10', clear
merge m:1 zcta10 using `ses2010', keep(master match) nogenerate
generate byte matched = !missing(AFFLUENCE)
summarize matched
local rows = r(N)
local matched = r(sum)
local rate = 100 * r(mean)
summarize AFFLUENCE
display "SES, ZCTA 2010 file: rows = `rows', matched = `matched', match rate = " ///
        %4.1f `rate' "%, mean affluence = " %6.3f r(mean)
tempfile vintage_2010
save `vintage_2010'
```

## Python

```{.python}
# Socioeconomic Status, ZCTA 2010 file (ICPSR 38528, DS0003). Its affluence
# measure is named for the ACS years behind it, so rename it to a common name
# now: the two files we are comparing call the same construct different things.
vintage_2010 = mydata_zcta10.merge(
    ses2008_2017_zcta10[["ZCTA10", "AFFLUENCE13_17"]].rename(columns={"AFFLUENCE13_17": "AFFLUENCE"}),
    how="left", left_on="zcta10", right_on="ZCTA10",
).drop(columns="ZCTA10")

def match_summary(df):
    return pd.Series({
        "rows":           len(df),
        "matched":        df["AFFLUENCE"].notna().sum(),
        "match_rate_pct": round(100 * df["AFFLUENCE"].notna().mean(), 1),
        "mean_affluence": round(df["AFFLUENCE"].mean(), 3),
    })

match_summary(vintage_2010)
```

:::

::: {.panel-tabset group="language"}
## R

```{r}
#| label: vintage-2020

# The same ZCTA codes — 2010-vintage codes, straight from the crosswalk above —
# against the ZCTA 2020 file (ICPSR 38528, DS0008). Nothing in this join knows
# that the codes and the geography come from different censuses. It runs, it
# matches, it returns a number.
vintage_2020 <- mydata_zcta10 %>%
  left_join(
    ses2018_2022_zcta20 %>% select(ZCTA20, AFFLUENCE),
    by = c("zcta10" = "ZCTA20")
  )

vintage_2020 %>%
  summarise(
    rows           = n(),
    matched        = sum(!is.na(AFFLUENCE)),
    match_rate_pct = round(100 * mean(!is.na(AFFLUENCE)), 1),
    mean_affluence = round(mean(AFFLUENCE, na.rm = TRUE), 3)
  )
```

## Stata

```{.stata}
* The same ZCTA codes, 2010-vintage codes straight from the crosswalk above,
* against the ZCTA 2020 file (ICPSR 38528, DS0008). Nothing in this merge knows
* that the codes and the geography come from different censuses. It runs, it
* matches, it returns a number.
use `mydata_zcta10', clear
merge m:1 zcta10 using `ses2020', keep(master match) nogenerate
generate byte matched = !missing(AFFLUENCE)
summarize matched
local rows = r(N)
local matched = r(sum)
local rate = 100 * r(mean)
summarize AFFLUENCE
display "SES, ZCTA 2020 file: rows = `rows', matched = `matched', match rate = " ///
        %4.1f `rate' "%, mean affluence = " %6.3f r(mean)
tempfile vintage_2020
save `vintage_2020'
```

## Python

```{.python}
# The same ZCTA codes, 2010-vintage codes straight from the crosswalk above,
# against the ZCTA 2020 file (ICPSR 38528, DS0008). Nothing in this join knows
# that the codes and the geography come from different censuses. It runs, it
# matches, it returns a number.
vintage_2020 = mydata_zcta10.merge(
    ses2018_2022_zcta20[["ZCTA20", "AFFLUENCE"]],
    how="left", left_on="zcta10", right_on="ZCTA20",
).drop(columns="ZCTA20")

match_summary(vintage_2020)
```

:::

::: {.panel-tabset group="language"}
## R

```{r}
#| label: vintage-compare

bind_rows(
  `SES, ZCTA 2010 file` = vintage_2010,
  `SES, ZCTA 2020 file` = vintage_2020,
  .id = "nanda_file"
) %>%
  group_by(nanda_file) %>%
  summarise(
    rows           = n(),
    matched        = sum(!is.na(AFFLUENCE)),
    match_rate_pct = round(100 * mean(!is.na(AFFLUENCE)), 1),
    mean_affluence = round(mean(AFFLUENCE, na.rm = TRUE), 3),
    .groups = "drop"
  )
```

## Stata

```{.stata}
use `vintage_2010', clear
generate str nanda_file = "SES, ZCTA 2010 file"
append using `vintage_2020'
replace nanda_file = "SES, ZCTA 2020 file" if missing(nanda_file)
generate byte one = 1
collapse (sum) rows = one (sum) matched (mean) match_rate_pct = matched ///
         (mean) mean_affluence = AFFLUENCE, by(nanda_file)
replace match_rate_pct = round(100 * match_rate_pct, 0.1)
replace mean_affluence = round(mean_affluence, 0.001)
list, clean noobs
```

## Python

```{.python}
pd.concat(
    {"SES, ZCTA 2010 file": vintage_2010, "SES, ZCTA 2020 file": vintage_2020},
    names=["nanda_file"],
).groupby(level="nanda_file").apply(match_summary)
```

:::

Read the table twice. The match rate is identical in both rows, so a match rate alone
would never have told you anything was wrong. The mean moved. That is the shape of a
vintage error: not a missing row you can count, but a plausible number attached to the
wrong place.

::: {.callout-warning}
## Pick your vintage from your data, not from the file that happens to be open

The year your participants were observed decides the vintage, not the file that happens
to be open. Observations through 2019 belong with 2010 geographies; from 2020 on, with
2020 geographies. Our synthetic file spans {{< var synthetic.years >}}, so it straddles
that boundary and also runs past the last year either NaNDA dataset covers. That is why
the SES step above split the file by observation year and joined each part to its own
file, and why Social Services was carried forward from its last available year with the
decision written down. The one thing that split did not do is convert the codes: the
2018-on half still carries 2010 ZCTA codes into a 2020 file. For a real analysis you
would run those rows through a 2020-target crosswalk first.
:::

## Who did you lose?

A match rate is a headline, not a diagnostic. The headline above tells you the size of
the loss and nothing about its shape. The diagnostic is whether the rows that
did not match differ from the rows that did.

In this file about {{< var synthetic.zip_missing_pct >}} of rows have no ZIP at all. If
those rows are disproportionately rural, or older, or in worse health, the matched sample
is no longer the sample you drew, and every neighborhood coefficient you estimate is
conditional on having a mailable address. That is a selection problem, and it does not
announce itself.

Neither language will do this for you. dplyr's joins are silent. Stata's `_merge`
reports the count, not the composition. The comparison has to be made deliberately:
isolate the unmatched rows, and put their means beside the matched rows' means on the
variables you actually care about.

::: {.panel-tabset group="language"}
## R

```{r}
#| label: match-rate

# The headline.
mydata_nanda %>%
  summarise(
    rows           = n(),
    matched        = sum(in_nanda),
    match_rate_pct = round(100 * mean(in_nanda), 1)
  )

# The same number, by year, which is where it stops being a headline. 2023
# onward matches because we carried Social Services forward; without that step
# these rows would show 0%.
mydata_nanda %>%
  group_by(year) %>%
  summarise(
    rows           = n(),
    matched        = sum(in_nanda),
    match_rate_pct = round(100 * mean(in_nanda), 1),
    .groups = "drop"
  )
```

## Stata

```{.stata}
* The headline.
use `mydata_nanda', clear
summarize in_nanda
display "rows = " r(N) ", matched = " r(sum) ", match rate = " %4.1f 100 * r(mean) "%"

* The same number, by year, which is where it stops being a headline. 2023
* onward matches because we carried Social Services forward; without that step
* these rows would show 0%.
tabulate year in_nanda, row nofreq
```

## Python

```{.python}
# The headline.
pd.Series({
    "rows":           len(mydata_nanda),
    "matched":        mydata_nanda["in_nanda"].sum(),
    "match_rate_pct": round(100 * mydata_nanda["in_nanda"].mean(), 1),
})

# The same number, by year, which is where it stops being a headline. 2023
# onward matches because we carried Social Services forward; without that step
# these rows would show 0%.
(mydata_nanda.groupby("year")["in_nanda"]
   .agg(rows="size", matched="sum", match_rate_pct=lambda s: round(100 * s.mean(), 1))
   .reset_index())
```

:::

::: {.panel-tabset group="language"}
## R

```{r}
#| label: anti-join

# anti_join() keeps the rows of the left table that found no partner — the
# complement of the join, and the half nobody looks at.
unmatched <- mydata_nanda %>%
  anti_join(
    socialservices_zcta10 %>% select(zcta10, year),
    by = c("zcta10", "year")
  )

# A count is not a diagnosis. A row can fail for four different reasons, and
# the reasons call for different answers: the first two are a data-collection
# problem, the third is a coverage limit you state in your methods, the fourth
# is usually a vintage mismatch. Three occur in this file; the third cannot,
# because we carried Social Services forward. It is named so the table says
# so if that step is ever dropped.
unmatched %>%
  mutate(reason = case_when(
    is.na(zip)    ~ "no ZIP to start from",
    is.na(zcta10) ~ "ZIP not in the crosswalk",
    year > 2022   ~ "year past NaNDA's coverage",
    TRUE          ~ "ZCTA not in the NaNDA file"
  )) %>%
  tabyl(reason)
```

## Stata

```{.stata}
* The rows that found no partner: the complement of the join, and the half
* nobody looks at. in_nanda == 0 is exactly the "ours only" side of the merge.
use `mydata_nanda', clear
keep if in_nanda == 0

* A count is not a diagnosis. A row can fail for four different reasons, and
* the reasons call for different answers: the first two are a data-collection
* problem, the third is a coverage limit you state in your methods, the fourth
* is usually a vintage mismatch. Three occur in this file; the third cannot,
* because we carried Social Services forward. It is named so the table says
* so if that step is ever dropped.
generate str reason = "ZCTA not in the NaNDA file"
replace  reason = "year past NaNDA's coverage" if year > 2022
replace  reason = "ZIP not in the crosswalk"   if missing(zcta10)
replace  reason = "no ZIP to start from"       if missing(zip)
tabulate reason
```

## Python

```{.python}
# The rows that found no partner: the complement of the join, and the half
# nobody looks at. indicator="left_only" is the anti-join.
unmatched = (
    mydata_nanda.merge(
        socialservices_zcta10[["zcta10", "year"]].drop_duplicates(),
        how="left", on=["zcta10", "year"], indicator=True,
    )
    .query('_merge == "left_only"')
    .drop(columns="_merge")
)

# A count is not a diagnosis. A row can fail for four different reasons, and
# the reasons call for different answers: the first two are a data-collection
# problem, the third is a coverage limit you state in your methods, the fourth
# is usually a vintage mismatch. Three occur in this file; the third cannot,
# because we carried Social Services forward. It is named so the table says
# so if that step is ever dropped.
unmatched.assign(reason=np.select(
    [unmatched["zip"].isna(), unmatched["zcta10"].isna(), unmatched["year"] > 2022],
    ["no ZIP to start from", "ZIP not in the crosswalk", "year past NaNDA's coverage"],
    default="ZCTA not in the NaNDA file",
))["reason"].value_counts()
```

:::

Three reasons appear, and they call for different responses. No ZIP is the big one and
a data-collection problem; nothing downstream fixes it. The other two turn up twice
each. A ZIP the crosswalk does not know is a typo or a ZIP newer than the crosswalk; a
ZCTA the NaNDA file does not carry usually means the two were built on different
vintages. At two rows apiece they vanish inside a match rate, and in real data they are
the ones telling you something upstream is misaligned. That is why the table lists
reasons, not just a count. The fourth reason, year past coverage, is absent only because
we carried Social Services forward. Had we not, 407 rows would sit in that line, and
the honest description of them is the same either way: a limit of the NaNDA file,
handled by a decision you state in your methods.

::: {.panel-tabset group="language"}
## R

```{r}
#| label: who-is-missing

# The question a match rate cannot answer: are the people who fell out different
# from the people who stayed, on the things the study is actually about?
mydata_nanda %>%
  group_by(in_nanda) %>%
  summarise(
    n                 = n(),
    loneliness        = round(mean(loneliness,        na.rm = TRUE), 2),
    physical_activity = round(mean(physical_activity, na.rm = TRUE), 2),
    own_pet           = round(mean(own_pet,           na.rm = TRUE), 3),
    .groups = "drop"
  )
```

## Stata

```{.stata}
* The question a match rate cannot answer: are the people who fell out different
* from the people who stayed, on the things the study is actually about?
use `mydata_nanda', clear
tabstat loneliness physical_activity own_pet, by(in_nanda) statistics(n mean) format(%9.3f)
```

## Python

```{.python}
# The question a match rate cannot answer: are the people who fell out different
# from the people who stayed, on the things the study is actually about?
(mydata_nanda.groupby("in_nanda")
   .agg(n=("in_nanda", "size"),
        loneliness=("loneliness", "mean"),
        physical_activity=("physical_activity", "mean"),
        own_pet=("own_pet", "mean"))
   .round({"loneliness": 2, "physical_activity": 2, "own_pet": 3})
   .reset_index())
```

:::

## A toy model, to show the join was for something

One model, not an analysis: does owning a pet go with the neighborhood measures we
just attached? The service count is cut into bins so the coefficients read as
comparisons against ZCTAs with none. Note the number of rows the model actually uses.
Every unmatched row drops out here, silently, which is what the diagnostics above were
for.

::: {.panel-tabset group="language"}
## R

```{r}
#| label: toy-model

mydata_nanda <- mydata_nanda %>%
  mutate(count_totindivfamilyservices_6cat = cut(
    count_totindivfamilyservices,
    breaks = c(-1, 0, 1, 2, 3, 5, 10, Inf),
    labels = c("0", "1", "2", "3", "4-5", "6-10", "11+"),
    right = TRUE
  ))

own_pet_model <- glm(
  own_pet ~ hot_meal_days + count_totindivfamilyservices_6cat + ANY_OPEN_PARK + AFFLUENCE,
  data = mydata_nanda, family = binomial
)
summary(own_pet_model)
```

## Stata

```{.stata}
use `mydata_nanda', clear
generate byte count_totindivfamilyservices_6cat = .
replace count_totindivfamilyservices_6cat = 1 if count_totindivfamilyservices == 0
replace count_totindivfamilyservices_6cat = 2 if count_totindivfamilyservices == 1
replace count_totindivfamilyservices_6cat = 3 if count_totindivfamilyservices == 2
replace count_totindivfamilyservices_6cat = 4 if count_totindivfamilyservices == 3
replace count_totindivfamilyservices_6cat = 5 if inrange(count_totindivfamilyservices, 4, 5)
replace count_totindivfamilyservices_6cat = 6 if inrange(count_totindivfamilyservices, 6, 10)
replace count_totindivfamilyservices_6cat = 7 if count_totindivfamilyservices >= 11 & !missing(count_totindivfamilyservices)
label define svc6 1 "0" 2 "1" 3 "2" 4 "3" 5 "4-5" 6 "6-10" 7 "11+"
label values count_totindivfamilyservices_6cat svc6

* ANY_OPEN_PARK is a labeled factor in the .dta; i. treats it as categorical.
logit own_pet hot_meal_days i.count_totindivfamilyservices_6cat i.ANY_OPEN_PARK AFFLUENCE
```

## Python

```{.python}
mydata_nanda["count_totindivfamilyservices_6cat"] = pd.cut(
    mydata_nanda["count_totindivfamilyservices"],
    bins=[-1, 0, 1, 2, 3, 5, 10, np.inf],
    labels=["0", "1", "2", "3", "4-5", "6-10", "11+"],
)

# pandas has no models. statsmodels is the one extra package this step needs
# (pip install statsmodels); everything above runs without it.
try:
    import statsmodels.formula.api as smf
except ImportError:
    smf = None
    print("statsmodels is not installed; skipping the toy model.")

if smf is not None:
    own_pet_model = smf.logit(
        "own_pet ~ hot_meal_days + C(count_totindivfamilyservices_6cat) + C(ANY_OPEN_PARK) + AFFLUENCE",
        data=mydata_nanda.dropna(subset=["count_totindivfamilyservices_6cat", "ANY_OPEN_PARK", "AFFLUENCE"]),
    ).fit()
    print(own_pet_model.summary())
```

:::

::: {.callout-tip}
## The four questions

1. What identifier do I actually have?
2. What vintage is it, and what vintage is the file?
3. What crosswalk, targeting what vintage?
4. What's my match rate, and are the unmatched rows different?

:::

---

# 4. Now what? {#sec-nowwhat}

*About {{< var timing.sec4 >}} minutes, then questions*


Everything above is the how. {{< var session.discussant >}} closes with a few minutes
on the why: what a linked file lets you ask that you could not ask before, a few
concrete examples of questions in aging research that neighborhood measures open up,
and where to start with NaNDA if one of them is yours. Two cautions come along for the
ride. A neighborhood measure describes a place, not a person, so be clear when it is
context and when it is being asked to stand in for exposure. And people in the same
ZCTA are not independent observations, so the analysis has to account for that.


Everything you saw today is already rendered on this page, so nothing depends on notes
taken during the session. The materials ZIP holds the synthetic data and its codebook,
the crosswalk, this walkthrough as a runnable `notebook.qmd`, the Stata and Python
equivalents in `stata/` and `python/`, and an offline copy of this page. The scripts
follow the same steps under the same section numbers, so you can read them beside this
walkthrough.

The NaNDA files are the one thing not in the ZIP. They come from ICPSR, which asks for a
free account, so you download your own — about five minutes, and `data/nanda/README.md`
in the ZIP walks you through it. That is only needed if you want to re-run the code.
If you use these materials in teaching or in a paper, cite them by the DOI below; it
always resolves to the latest version.

## Take this with you

| | |
|---|---|
| This walkthrough | [{{< var site.url >}}notebook.html]({{< var site.url >}}notebook.html) |
| Everything, zipped | [{{< var site.zip >}}]({{< var site.zip >}}) |
| Source | [{{< var site.repo >}}]({{< var site.repo >}}) |
| Cite it | {{< var site.doi >}} |

## Where to go next

- **The NaNDA catalogue on openICPSR**, <https://www.openicpsr.org/openicpsr/nanda>.
  Every dataset, every geography, free to download.
- **The NaNDA website**, <https://nanda.isr.umich.edu/>. Documentation, news, and how to
  reach the team.
- **Our published merge code**: {{< var crosswalk.merge_code_name >}}
  ({{< var crosswalk.merge_code_year >}}), <{{< var crosswalk.merge_code_doi >}}>.
- **The crosswalk**: {{< var crosswalk.primary_name >}},
  <{{< var crosswalk.primary_url >}}>.
- **Questions after today**: <nanda-admin@umich.edu>


---

## Session information {.appendix}

::: {.panel-tabset group="language"}
## R

```{r}
#| label: session-info

sessionInfo()
```

## Stata

```{.stata}
about
```

## Python

```{.python}
import sys
print(sys.version)
print("pandas", pd.__version__, "| numpy", np.__version__)
```

:::
